---
title: "Undirected Spectral Clustering Examples"
author: "Carlyle Morgan"
date: "5/24/2021"
output: pdf_document
---

```{r setup, results='hide', message=FALSE}
library(igraph)
library(dplyr)
library(stringr)
library(readr)
library(stats)
library(expm)
```


## Introduction

This dataset illustrates collaboration efforts between various academics in undirected network form. There are 86 distinct authors and 124 edges, each weighted by the number of collaborations between each of the authors.

## Preprocessing
Importing the dataset and reordering the to and from columns:
```{r, message=FALSE}
authors <- read_table2("C:/Users/sc_mo/Documents/R/Projects/Stats Research/ca-sandi_auths.csv")
auth_temp<-authors$node1
authors$node1<-authors$node2
authors$node2<-auth_temp
authlist<- unique(c(authors$node1, authors$node2))
```

Calling the network data into igraph and applying the edge weights as provided by the dataset:
```{r, message=FALSE}
g.authors<- graph.data.frame(authors[1:2], vertices = authlist, directed = FALSE)
E(g.authors)$weight<-authors$weight
set.seed(123)
plot(g.authors)
```

Checking for connected components and simplicity:
```{r, message=FALSE}
is.simple(g.authors)
is.connected(g.authors)
```



## Creating the Necessary Laplacian Components



Creating the weighted(W) and unweighted(A) adjacency matrices:
```{r, message=FALSE}
A<-get.adjacency(g.authors)
A<-as.matrix(A)
W<-A
for (i in 1:nrow(authors)){
  W[as.character(authors$node1[i]), as.character(authors$node2[i])]<-authors$weight[i]
  W[as.character(authors$node2[i]), as.character(authors$node1[i])]<-authors$weight[i]
}
```

Creating the degree matrix(D):
```{r, message=FALSE}
D<-diag(strength(g.authors))
```

## Generating the Laplacian Matrices:

Creating the unnormalized Laplacian Matrix, $L_{un} = D - W$:
```{r, message=FALSE}
L_unnorm<-D-W
```

Creating the normalized symmetric Laplacian Matrix, $L = I - D^{-1/2}WD^{-1/2}$:
```{r, message=FALSE}
L_normsym<-D-W
```

Creating the normalized random walk Laplacian Matrix, $ L = I - D^{-1}W$
```{r, message=FALSE}
L_normrw<-D-W
```

## Unnormalized Spectral Clustering

Calculating the eigenvectors of $L_{un}$:
```{r}
evL <- eigen(L_unnorm, symmetric=TRUE)
```

Creating a simple plot of the eigenvalues:
```{r, message=FALSE, warning=FALSE}
plot(1:10,rev(evL$values)[1:10], log="y")
abline(v=3.5, col="red", lty=3)
abline(v=6.5, col="blue", lty=2)
```

### Clustering with k=2 clusters:
```{r}
k   <- 2
Z   <- evL$vectors[,(ncol(evL$vectors)-k+1):ncol(evL$vectors)]

km <- kmeans(Z, centers=k, nstart=5)
V(g.authors)$cluster<-km$cluster
V(g.authors)[cluster == 1]$color<-"red"
V(g.authors)[cluster == 2]$color<-"dodgerblue"
set.seed(123)
plot(g.authors)
```

### Clustering with k=3 clusters:
```{r}
k   <- 3
Z   <- evL$vectors[,(ncol(evL$vectors)-k+1):ncol(evL$vectors)]

km <- kmeans(Z, centers=k, nstart=5)
V(g.authors)$cluster<-km$cluster
V(g.authors)[cluster == 1]$color<-"red"
V(g.authors)[cluster == 2]$color<-"dodgerblue"
V(g.authors)[cluster == 3]$color<-"yellow"
set.seed(123)
plot(g.authors)
```